---
title: "Belief Networks grouped by ideology (5-scale)"
output:
  html_document:
    df_print: paged
---

# Preliminaries

```{r}
library(here)
knitr::opts_knit$set(root.dir = here())
knitr::opts_chunk$set(dev = "png", fig.align = "center",
                      fig.width = 12, fig.height = 8, fig.path = "../figures/",
                      echo = FALSE, warning = FALSE, message = FALSE, error = FALSE)
```


```{r}
setwd(here())
source("code/data-wrangling.R")
```

```{r}
library(psychonetrics)
```

```{r}
model_data <- data %>% select(c(network_variables,lr5))
model_data <- rescale_data(model_data)
model_data <- model_data |> filter(!is.na(model_data$lr5))
```

## Specification of the models

Models:

1.  Empty model, all parameters free `model_1`
2.  Sparse model, all parameters free `model_1b`
3.  Empty model, equal networks `model_2`
4.  Sparse model, equal networks `model_2b`
5.  Empty model, equal temperature `model_3`
6.  Sparse model, equal temperature `model_3b`
7.  Empty model, equal networks and temperature `model_4`
8.  Sparse model, equal networks and temperature `model_4b`

```{r cache=TRUE}
model_1 <- ggm(model_data ,omega = "empty",group = "lr5")

model_1 <- model_1 %>% runmodel

model_1b <- model_1  %>%  stepup(
        criterion = "bic",
        alpha = 0.00001,
        greedy = FALSE,
        greedyadjust = "fdr"
    )

model_2 <- model_1 %>% groupequal("omega") %>% runmodel

model_2b <- model_2 %>%  stepup(
    criterion = "bic",
    mi = "mi_equal",
    alpha = 0.00001,
    greedy = FALSE,
    greedyadjust = "fdr"
)

model_3 <- model_1 %>% groupequal("delta") %>% runmodel

model_3b <- model_3 %>%  stepup(
    criterion = "bic",
    mi = "mi_equal",
    alpha = 0.00001,
    greedy = FALSE,
    greedyadjust = "fdr"
)

model_4 <- model_2 %>% groupequal("delta") %>% runmodel

model_4b <- model_4 %>%  stepup(
    criterion = "bic",
    mi = "mi_equal",
    alpha = 0.00001,
    greedy = FALSE,
    greedyadjust = "fdr"
)

```

```{r}
comparison <- compare(
    `1. all parameters free (empty)` = model_1,
    `2. all parameters free (sparse)` = model_1b,
    `3. equal networks (empty)` = model_2,
    `4. equal networks (sparse)` = model_2b,
    `5. equal temperature (empty)` = model_3,
    `6. equal temperature (sparse)` = model_3b,
    `7. equal networks & temperature (empty)` = model_4,
    `8. equal networks & temperature (sparse)` = model_4b) %>% arrange(BIC) 

comparison <- as.data.frame(comparison)

rownames(comparison) <- NULL

comparison |> select(model,BIC)
```
```{r}
setwd(here())
saveRDS(getmatrix(model_1b, "omega"),file='data/lr_matrices.Rds')
```


## Model temperature

```{r}
delta_matrix <-  getmatrix(model_1b, "delta")
deltas <-
    c(
        sum(diag(delta_matrix$FL)) / nrow(delta_matrix$FL),
        sum(diag(delta_matrix$L)) / nrow(delta_matrix$L),
        sum(diag(delta_matrix$C)) / nrow(delta_matrix$C),
        sum(diag(delta_matrix$R)) / nrow(delta_matrix$R),
        sum(diag(delta_matrix$FR)) / nrow(delta_matrix$FR)
        
    )
SEs <- model_1b@parameters$se[model_1b@parameters$matrix == "delta"]

se <- c(mean(SEs[1:7]),mean(SEs[8:14]),mean(SEs[15:21]),mean(SEs[22:28]),mean(SEs[29:35]))

df <- data.frame(
    temperature = deltas,
    group = c("Far Left","Left","Center","Right","Far Right"),
    lower = (deltas-qnorm(0.975) * se),
    upper = (deltas+qnorm(0.975) * se),
    stringsAsFactors = FALSE
)
```

```{r lr5-model1b-temperature}
df$fixed <- is.na(df$lower)
df$group <- factor(df$group, levels = c("Far Left","Left","Center","Right","Far Right"))

g <- ggplot(df,aes(x=as.numeric(group), y = temperature, ymin = lower, ymax = upper)) +
    geom_line() +
    geom_errorbar(width = 0.05) +
    geom_point(cex = 3, colour = "black") +
    geom_point(aes(colour = fixed), cex = 3) +  theme_bw()  +
    xlab("") + ylab(expression(paste("Temperature"))) +
    scale_x_continuous(breaks =  1:5, labels = levels(df$group), expand = c(0.1,0.1)) +
    scale_y_continuous( limits = c(0.3,0.45)) +
    theme( panel.grid.major.x = element_blank(),panel.grid.minor.x = element_blank())+
    ggtitle(expression(paste("Model 2: 3 groups; ",bold(Omega)," sparse and free; ",bold(tau)," free; "))) +
    scale_colour_manual(values = c("black","white")) +
    theme(legend.position = "none")

print(g)
```


```{r}
df[,1:4]
```

## Influence

```{r}
setwd(here())
source("code/k-shell-influence.R")
library(psych)
```

```{r}
columns = ncol(model_data) - 1

factor_levels <- 5
```

```{r}
fl_matrix <- getmatrix(model_1b,matrix = "omega",group="FL")
graph <-  graph_from_adjacency_matrix(fl_matrix,
                                    weighted = TRUE,
                                    mode = "undirected")
V(graph)$label <- as.character(1:7)
  
results <- get_influence(graph)
results$group = "FL"
```


```{r}
l_matrix <- getmatrix(model_1b,matrix = "omega",group="L")
graph <-  graph_from_adjacency_matrix(l_matrix,
                                    weighted = TRUE,
                                    mode = "undirected")
V(graph)$label <- as.character(1:7)
  
result <- get_influence(graph)
result$group <- "L"
results <- rbind(results,result)
```

```{r}
c_matrix <- getmatrix(model_1b,matrix = "omega",group="C")
graph <-  graph_from_adjacency_matrix(c_matrix,
                                    weighted = TRUE,
                                    mode = "undirected")
V(graph)$label <- as.character(1:7)
  
result <- get_influence(graph)
result$group <- "C"
results <- rbind(results,result)

```

```{r}
r_matrix <- getmatrix(model_1b,matrix = "omega",group = "R")
graph <-  graph_from_adjacency_matrix(r_matrix,
                                    weighted = TRUE,
                                    mode = "undirected")
V(graph)$label <- as.character(1:7)
  
result <- get_influence(graph)
result$group <- "R"
results <- rbind(results,result)

```

```{r}
fr_matrix <- getmatrix(model_1b,matrix = "omega",group = "FR")
graph <-  graph_from_adjacency_matrix(fr_matrix,
                                    weighted = TRUE,
                                    mode = "undirected")
graph_vertices <- V(graph)
  
V(graph)$label <- as.character(1:7)
  
result <- get_influence(graph)
result$group <- "FR"
results <- rbind(results,result)

```

```{r}
lr5_inf <- as.data.frame(results)
lr5_inf
```

```{r}
e = array()
gic_means = e 
gi_means = e
gsm_means = e
gic_sds = e
gi_sds = e
gsm_sds = e
gic_k = e 
gi_k = e
gsm_k = e

d <- lr5_inf |> filter(group == "FL")

gic_means[1] = mean(unlist(d[1,1:7]))
gi_means[1] = mean(unlist(d[3,1:7]))
gsm_means[1] = mean(unlist(d[4,1:7]))
gic_sds[1] = sd(unlist(d[1,1:7]))
gi_sds[1] = sd(unlist(d[3,1:7]))
gsm_sds[1] = sd(unlist(d[4,1:7]))
gic_k[1] = kurtosi(unlist(d[1,1:7]))
gi_k[1] = kurtosi(unlist(d[3,1:7]))
gsm_k[1] = kurtosi(unlist(d[4,1:7]))
```

```{r}
d <- lr5_inf |> filter(group == "L")

gic_means[2] = mean(unlist(d[1,1:7]))
gi_means[2] = mean(unlist(d[3,1:7]))
gsm_means[2] = mean(unlist(d[4,1:7]))
gic_sds[2] = sd(unlist(d[1,1:7]))
gi_sds[2] = sd(unlist(d[3,1:7]))
gsm_sds[2] = sd(unlist(d[4,1:7]))
gic_k[2] = kurtosi(unlist(d[1,1:7]))
gi_k[2] = kurtosi(unlist(d[3,1:7]))
gsm_k[2] = kurtosi(unlist(d[4,1:7]))
```

```{r}
d <- lr5_inf |> filter(group == "C")

gic_means[3] = mean(unlist(d[1,1:7]))
gi_means[3] = mean(unlist(d[3,1:7]))
gsm_means[3] = mean(unlist(d[4,1:7]))
gic_sds[3] = sd(unlist(d[1,1:7]))
gi_sds[3] = sd(unlist(d[3,1:7]))
gsm_sds[3] = sd(unlist(d[4,1:7]))
gic_k[3] = kurtosi(unlist(d[1,1:7]))
gi_k[3] = kurtosi(unlist(d[3,1:7]))
gsm_k[3] = kurtosi(unlist(d[4,1:7]))
```

```{r}
d <- lr5_inf |> filter(group == "R")

gic_means[4] = mean(unlist(d[1,1:7]))
gi_means[4] = mean(unlist(d[3,1:7]))
gsm_means[4] = mean(unlist(d[4,1:7]))
gic_sds[4] = sd(unlist(d[1,1:7]))
gi_sds[4] = sd(unlist(d[3,1:7]))
gsm_sds[4] = sd(unlist(d[4,1:7]))
gic_k[4] = kurtosi(unlist(d[1,1:7]))
gi_k[4] = kurtosi(unlist(d[3,1:7]))
gsm_k[4] = kurtosi(unlist(d[4,1:7]))
```

```{r}
d <- lr5_inf |> filter(group == "FR")

gic_means[5] = mean(unlist(d[1,1:7]))
gi_means[5] = mean(unlist(d[3,1:7]))
gsm_means[5] = mean(unlist(d[4,1:7]))
gic_sds[5] = sd(unlist(d[1,1:7]))
gi_sds[5] = sd(unlist(d[3,1:7]))
gsm_sds[5] = sd(unlist(d[4,1:7]))
gic_k[5] = kurtosi(unlist(d[1,1:7]))
gi_k[5] = kurtosi(unlist(d[3,1:7]))
gsm_k[5] = kurtosi(unlist(d[4,1:7]))
```


```{r}
lr5_summary <- cbind(df[,1],gic_means,gic_sds,gic_k,gi_means,gi_sds,gi_k,gsm_means,gsm_sds,gsm_k)
colnames(lr5_summary) <- c("Temperature","GIC Mean","GIC SD","GIC Kurtosis","GI Mean","GI SD","GI Kurtosis", "GSM Mean ","GSM SD","GSM Kurtosis")
rownames(lr5_summary) <- c("Far Left","Left","Center","Right","Far Right")
lr5_summary <- round(lr5_summary,2)
lr5_summary
```

## Energy

```{r}
group_data <- model_data |> filter(lr5 == "FL") |>  drop_na()
n = nrow(group_data)
res = matrix(0,7,7)
```

```{r cache=TRUE}
for (i in 1:n) {
  for (j in 1:6) {
    for (k in (j + 1):7) {
      res[j,k] = res[j,k] + fl_matrix[j,k]*as.numeric(group_data[i,j])*as.numeric(group_data[i,k])
    }
  }
}
```

```{r}
lr5_fl_e <- -sum(res[upper.tri(res)]/n)
```


```{r}
group_data <- model_data |> filter(lr5 == "L") |>  drop_na()
n = nrow(group_data)
res = matrix(0,7,7)
```

```{r cache=TRUE}
for (i in 1:n) {
  for (j in 1:6) {
    for (k in (j + 1):7) {
      res[j,k] = res[j,k] + l_matrix[j,k]*as.numeric(group_data[i,j])*as.numeric(group_data[i,k])
    }
  }
}
```

```{r}
lr5_l_e <- -sum(res[upper.tri(res)]/n)
```

```{r}
group_data <- model_data |> filter(lr5 == "C") |>  drop_na()
n = nrow(group_data)
res = matrix(0,7,7)
```

```{r cache=TRUE}
for (i in 1:n) {
  for (j in 1:6) {
    for (k in (j + 1):7) {
      res[j,k] = res[j,k] + c_matrix[j,k]*as.numeric(group_data[i,j])*as.numeric(group_data[i,k])
    }
  }
}
```

```{r}
lr5_c_e <- -sum(res[upper.tri(res)]/n)
```

```{r}
group_data <- model_data |> filter(lr5 == "R") |>  drop_na()
n = nrow(group_data)
res = matrix(0,7,7)
```

```{r cache=TRUE}
for (i in 1:n) {
  for (j in 1:6) {
    for (k in (j + 1):7) {
      res[j,k] = res[j,k] + r_matrix[j,k]*as.numeric(group_data[i,j])*as.numeric(group_data[i,k])
    }
  }
}
```

```{r}
lr5_r_e <- -sum(res[upper.tri(res)]/n)
```

```{r}
group_data <- model_data |> filter(lr5 == "FR") |>  drop_na()
n = nrow(group_data)
res = matrix(0,7,7)
```

```{r cache=TRUE}
for (i in 1:n) {
  for (j in 1:6) {
    for (k in (j + 1):7) {
      res[j,k] = res[j,k] + fr_matrix[j,k]*as.numeric(group_data[i,j])*as.numeric(group_data[i,k])
    }
  }
}
```

```{r}
lr5_fr_e <- -sum(res[upper.tri(res)]/n)
```

```{r}
lr5_summary <- cbind(lr5_summary,c(lr5_fl_e,lr5_l_e,lr5_c_e,lr5_r_e,lr5_fr_e))
colnames(lr5_summary) <- c("Temperature","GIC Mean","GIC SD","GIC Kurtosis","GI Mean","GI SD","GI Kurtosis", "GSM Mean ","GSM SD","GSM Kurtosis","Avg. Energy")
```

```{r}
lr5_summary 
```

```{r}
df$energy <- unlist(-lr5_summary["Average Energy"])
```

```{r lr5-model1b-temp-energy}
g <- ggplot(df,aes(x=as.numeric(group))) +
    geom_line(aes(y = temperature, colour = "#FC4E07")) +
    geom_point(aes(y = temperature),cex = 3, colour = "black") +
    geom_line(aes(y = energy,colour = "#00AFBB"), ) +
    geom_point(aes(y = energy),cex = 3, colour = "black") + theme_bw()  +
    xlab("") +
    scale_x_continuous(breaks =  1:5, labels = levels(df$group), expand = c(0.1,0.1)) +
    scale_y_continuous( limits = c(0.2,0.6)) +
    theme( panel.grid.major.x = element_blank(),panel.grid.minor.x = element_blank())+
    ggtitle(expression(paste("Model 2: 3 groups; ",bold(Omega)," sparse and free; ",bold(tau)," free; "))) +
    theme(legend.position = "right") +
     scale_color_identity(name = "",labels = c( " Negative of \n Avg. Energy"," Temperature"), guide= "legend")
print(g)
```

```{r}
setwd(here())
saveRDS(lr5_summary,file = "data/lr5_summary.Rds")
saveRDS(results,file="data/lr5_influence.Rds")
```
